---
title: "R-code used in Part I"
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: "Julian Quandt"
output: html_document
highlight: github
---


```{r setup, include=FALSE}
require(knitr)
require(BayesFactor)
knitr::opts_chunk$set(echo =T, warning = F, cache = F)
options(width = 80)
options(scipen = 10)
devAskNewPage(ask = FALSE)
```

# Slide 10 - Bayes Factor Calculation 

To calculate a Bayes Factor in this toy example (we would never really do it like this but use the `BayesFactor` package instead), we calculate the probability that we observe 60 women out of 100 people under h0 (first line of code), then we calculate it for H1 by calculating its probability for anything between 0 and 1. 
Then, we marignalize the porbabilities for the alternative by dividing the sum of these probabilities by the number of probabilities that we tried.
Third, we compare these probabilities.

```{r bf_calc}
h0 <- dbinom(60,100,0.5)  # get the probability of the data (i.e. 60 women) under H0
h1 <- dbinom(60, 100, seq(0,1, length.out = 1e6))  # probability of the data under H1 = uniform prior
h1_p <- sum(h1)/length(h1)  # marginalize the probabilities under H1
h0/h1_p # Bayes Factor of H0 over H1
```


# Slide 12 - sequential plots for p-value and Bayes Factor if H0 is true

```{r plot_h0}

set.seed(111) # make plot reproducible
g1 <- c() # create group 1
g2 <- c() # create group 2
pval_vec <- c() # collection of p-values
bf_vec <- c() # collection of BFs
for(i in 1:1000){
  g1_add <- rnorm(10,10,5) # add 10 participants to group 1
  g2_add <- rnorm(10,10,5) # add 10 participants to group 2
  g1 <- append(g1, g1_add)
  g2 <- append(g2, g2_add)
  pval <- t.test(g1,g2,paired = F, equal.var = T)$p.value # get p-value for current sample and store
  pval_vec <- append(pval_vec, pval) 
  bfac <- exp(ttestBF(g1,g2, paired = F)@bayesFactor$bf)# get BF for current sample and store (is returned as log-bf therefore exponentiate)
  bf_vec <- append(bf_vec, bfac)
}


n_test <- seq(10,1e4, by = 10) # x-axis for plot

# plot p-values
plot(n_test, pval_vec, main = "p-value if H0 is true", xlab = "sample size", ylab = "p-value")
abline(h = .05)

# plot BF
bf_vec_plot <- ifelse(bf_vec < 1, (1/bf_vec), -bf_vec) # convert BFs in favor of H1 from fractions to negative numbers
plot(n_test, bf_vec_plot, main = "Bayes Factor if H0 is true", xlab = "sample size", ylab = "Bayes Factor in favor of H0")
abline(h = 3, col = "green", lwd = 2)
abline(h = -3, col = "red", lwd = 2)

```

# Slide 13 - seqeuntial plots for p-value and Bayes Factor for d = .20

```{r plot_h1}

set.seed(999) # make plot reproducible

g3 <- c() # create group 1
g4 <- c() # create group 2
pval_vec2 <- c()
bf_vec2 <- c()
for(i in 1:1000){
  g3_add <- rnorm(10,11,5) # add 10 participants to group 1
  g4_add <- rnorm(10,10,5) # add 10 participants to group 2
  g3 <- append(g3, g3_add)
  g4 <- append(g4, g4_add)
  pval <- t.test(g3,g4,paired = F, equal.var = T)$p.value # get p-value for current sample and store
  pval_vec2 <- append(pval_vec2, pval)
  bfac <- exp(ttestBF(g3,g4, paired = F)@bayesFactor$bf) # get BF for current sample and store (is returned as log-bf therefore exponentiate)
  bf_vec2 <- append(bf_vec2, bfac)
}

bf_vec2_plot <- ifelse(bf_vec2 < 1, -(1/bf_vec2), bf_vec2)

# plot p-values
plot(n_test[1:50], pval_vec2[1:50], main = "p-value for small effect of d = .20", xlab = "sample size", ylab = "p-value")
abline(h = .05)

# plot BF
plot(n_test[1:50], bf_vec2_plot[1:50], main = "Bayes Factor for small effect of d = .20", xlab = "sample size", ylab = "Bayes Factor")
abline(h = 3, col = "green", lwd = 2)
abline(h = -3, col = "red", lwd = 2)
```

